Non-Nested Two Phase Debiased Calibration Simulation

### 2024-04-29 final/reports/prelim_paper.tex
Non-nested simulation working for HT and regression estimators.

where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
%\textcolor{red}{I don't know if it makes sense to change the Phase 2
%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
%I do, how is the two phase setup useful?} 
%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
%  $A_c$. The second-phase sample would be $A_2$. 
%}
We can combine these estimates to get
$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
a regression estimator as

@@

where % \textcolor{red}{(we may need transpose in $W$)}

@@

$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$

>>>>

where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
%\textcolor{red}{I don't know if it makes sense to change the Phase 2
%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
%I do, how is the two phase setup useful?} 
%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
%  $A_c$. The second-phase sample would be $A_2$. 
%}
We can combine these estimates using the effective sample size 
(\cite{kish1965survey}) to get
$\hat X_c = (n_{1, eff} \hat X_1 + n_{2, eff}\hat X_2) / (n_{1, eff} + n_{2,
eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective sample size for
$A_1$ and $A_2$ respectively. Then we can define a regression estimator as

@@

where, $W = n_{1, eff} / (n_{1, eff} + n_{2, eff})$, and

@@

$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W V(\hat X_1) W \beta_q^*$$

### 2024-04-29 src/explore/20240425-nndcsim.qmd
Non-nested simulation working for HT and regression estimators.

# source("GEcalib_code.R")

@@

@@

gen_samps(upop_df, "srs", "poisson")

@@

>>>>

@@

ht_est <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {

  theta <- sum(p2_df$Y / p2_df$pi2) / pop_obs

  if (sampling == "srs-poisson") {

    v_hat <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / pop_obs^2

  } else {

    v_hat <- NA

  }

  return(list(theta = theta, v_hat = v_hat))
}

reg_est <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {

  v_poisson <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / pop_obs^2
  v_srs <- (1 / nrow(p2_df) - 1 / pop_obs) * var(p2_df$Y)
  d_eff <- v_poisson / v_srs
  n2_eff <- nrow(p2_df) / d_eff
  n1_eff <- nrow(p1_df)

  x1_hat <- 
    c(sum(p1_df$X1 / p1_df$pi1) / pop_obs,
      sum(p1_df$X2 / p1_df$pi1) / pop_obs)

  x2_hat <- 
    c(sum(p2_df$X1 / p2_df$pi2) / pop_obs,
      sum(p2_df$X2 / p2_df$pi2) / pop_obs)

  w <- n1_eff / (n1_eff + n2_eff)
  xc_hat <- (n1_eff * x1_hat + n2_eff * x2_hat) / (n1_eff + n2_eff)

  # Estimate beta
  mod <- lm(Y ~ X1 + X2, data = p2_df)

  # Regression Estimation
  y_ht <- sum(p2_df$Y / p2_df$pi2) / pop_obs
  theta <- y_ht + sum((xc_hat - x2_hat) * mod$coefficients[2:3])

  # Variance Estimation
  coefs <- mod$coefficients
  eps <- (p2_df$Y - coefs[1] * w - coefs[2] * w * p2_df$X1 - coefs[3] * w * p2_df$X2)
  var_x <- 
  (1 / nrow(p1_df) - 1 / pop_obs) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))

  v1 <- w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
  v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / pop_obs^2

  cov_xy <- c(sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[2] * 
                 (n2_eff) / (n1_eff + n2_eff) * p2_df$X1 * eps) / pop_obs^2,
              sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[3] * 
                 (n2_eff) / (n1_eff + n2_eff) * p2_df$X3 * eps) / pop_obs^2)
  v3 <- sum(cov_xy)

  v_hat <- v1 + v2 + v3

  return(list(theta = theta, v_hat = as.numeric(v_hat)))
}

@@

clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
registerDoParallel(clust)

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "nleqslv", "rlang")) %dopar% {

    if (iter %% 100 == 0) {
      print(paste0("Iter: ", iter))
    }

    set.seed(iter)

    samps <- gen_samps(upop_df, "srs", "poisson")
    p1_df <- samps[[1]]
    p2_df <- samps[[2]]

    ht <- ht_est(p1_df, p2_df, nrow(upop_df))
    reg <- reg_est(p1_df, p2_df, nrow(upop_df))

    return(
      tibble(Est = c("HT", "Reg"),
             Theta = c(ht[[1]], reg[[1]]),
             Var = c(ht[[2]], reg[[2]]),
             Iter = iter)
    )
  } %>% bind_rows()

stopCluster(clust)

@@

```{r}
#| label: analyze results

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest)

%>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest)

%>%
  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%
  cat(file = "tables/nndcsim_mean.tex")

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims))

%>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, MCVar, EstVar, VarVar, Ttest)

%>%
  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 4) %>%
  cat(file = "tables/nndcsim_var.tex")
